\Large \bf Cyclone: Replication Middleware for NVM Applications

### 2016-09-22 paper/paper.tex
More stuff in the paper.

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}

\subsection*{Abstract}
Memory technology in the datacenter is due to undergo a paradigm shift with the
increased use of directly addressable non-volatile memory both in the form of
battery backed non-volatile DIMMs as well as newer memory types such as
3D XPoint. In anticipation, libraries that allow programmers to build durable
data structures on non-volatile heaps have become available. A missing piece
however is a component to make such durable data structures available using
replication across a commodity network such as ethernet. Existing solutions that
offer good performance are biased towards pure transactions across RDMA with a
dependence on external services and application specific code for fault recovery.

Cyclone is replication middleware that aims to add fault tolerance as
transparently as possible. A design for Cyclone is that NVM application
developers should not need to add fault recovery code. A second design goal for
Cyclone is to strive for good performance on commodity ethernet - with
innovations in the software stack rather than a requirement for RDMA capable
networking hardware.

Cyclone achieves these goals by providing a simple API for client server
applications that allows server state to be replicated by executing the same
sequence of RPC calls across all replicas. It provides at most once semantics
with automatic failover to hide fault recovery from programmers. Finally it uses
DPDK - a high performance, low latency network stack coupled with batching,
zero-copy and the ability to scale replication performance across
multiple instances of a consensus protocol to replicate millions of small
payload RPC messages a second on commodity ethernet.

\section{Introduction}
TBD
\section{Single Node Programming Model}
Cyclone starts with the assumption that the programmer has already built a
concurrent durable application on a non-volatile heap. A number of libraries and
filesystems make this possible today [DAX, NVML, others ?]. As a starting point,
we support applications built on top of Intel's Non Volatile Memory Library
(NVML) [nvml.io]. We describe NVML's programming model in this section to
provide adequate background for the rest of this paper. The primary feature of
interest is the ability to write code that is crash consistent. In other words
code that makes changes to the non-volatile heap is enclosed in a crash
consistent transaction such that either all of the transaction executes or none
of it executes in the face of power failures. This is critical to maintaining
consistency as the data structure, being durable, must survive across power
failures. To illustrate this idea, Figure~\ref{fig:example} shows how one might
build a persistent linked list using NVML - the key idea is to enclose the
entire process of updating the linked list within a crash consistent
transaction. This ensure that the linked list is not corrupted due to a power
failure. A second point worth noting is that NVM libraries such as NVML usually
provide a way to wrap pointers to non-volatile memory (using {\tt TOID} macros)
in order that such memory may be mapped to different locations in virtual memory
across restarts without impact to the accessing application.

@@

\section{Evaluation}

@@

\section{Conclusion}

\end{document}

>>>>

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
%\thispagestyle{empty}

\subsection*{Abstract}
The introduction of non-volatile directly attached memory in the datacenter - in
the form of NVDIMMs and new memory technologies allows efficient datastructures
to be built on non-volatile heaps that provide good performance coupled with
durability. However availability remains a problem for these single machine
solutions with proposed solutions counting on transactional interfaces built
over RDMA capable networking hardware that is often unavailable in datacenters.
Cyclone is designed to help solve this problem. It works as a transparent layer
with Intel's Non Volaile Memory Library (NVML) and allows the construction of
durable data-structures targeted towards mapping keys to values. We show that
Cyclone can replicate as many as 8 million updates a second using just 4
commodity 10 GigE ports on a machine. Cyclone is designed around a userspace
networking stack to reduce latency and can scale its consensus protocol across
cores and network ports to provide good throughput under load.

\section{Introduction}
Memory technology in the datacenter is due to undergo a paradigm shift with the
increased use of directly addressable non-volatile memory both in the form of
battery backed non-volatile DIMMs as well as newer memory types such as 3D
XPoint. In anticipation, libraries that allow programmers to build durable data
structures on non-volatile heaps have become available - an example of a such a
library that we have used in this paper is Intel's non volatile memory library
(NVML). A missing piece however is a component to make such durable data
structures available using replication across a commodity network such as
ethernet. Existing solutions that offer good performance are biased towards pure
transactions across RDMA with a dependence on external services and application
specific code for fault recovery [FARM, etc]. This is a problem for programmers
using NVML to construct durable and available datastructures where one cannot
apriori guarantee the availability of RDMA in the target environment especially
when running on externally hosted environments such as the cloud.

Cyclone is a prototype replication middleware that aims to demonstrate that one
can add fault tolerance transparently to a specific class of client server
applications written on top of libraries such as NVML. We specifically target
key-value maps in this paper and in that context make two key
contributions. First, we show that a simple API allows server state to be
replicated without the programmer needing to worry about fault recovery. This
allows one to start by building a concurrent durable data structure using
libraries such as NVML and add availability as a later step using Cyclone. The
second contribution we make is to show that one can achieve reasonably good
replication performance using commodity ethernet via innovations in the
networking stack and by scaling replication with the application. This result
should be a useful design point for developers looking to deploy applications on
non-volatile heaps in commodity environments such as public clouds.

\section{Single Node Programming Model}
Cyclone starts with the assumption that the programmer has already built a
concurrent durable application on a non-volatile heap. A number of libraries and
filesystems make this possible today [DAX, NVML, others ?]. As a starting point,
we support applications built on top of Intel's Non Volatile Memory Library
(NVML) [nvml.io]. We describe NVML's programming model in this section to
provide adequate background for the rest of this paper. The primary feature of
interest is the ability to write code that is crash consistent. In other words
code that makes changes to the non-volatile heap is enclosed in a crash
consistent transaction such that either all of the transaction executes or none
of it executes in the face of power failures. This is critical to maintaining
consistency as the data structure, being durable, must survive across power
failures. To illustrate this idea, Figure~\ref{fig:example} shows how one might
build a persistent linked list using NVML - the key idea is to enclose the
entire process of updating the linked list within a crash consistent
transaction. This ensure that the linked list is not corrupted due to a power
failure. A second point worth noting is that NVM libraries such as NVML usually
provide a way to wrap pointers to non-volatile memory (using {\tt TOID} macros)
in order that such memory may be mapped to different locations in virtual memory
across restarts without impact to the accessing application.

@@

\section{Evaluation}
We evaluate Cyclone against its design goals by using it for up to 3-way
replication in a cluster of twelve 48 core Xeon servers. Three servers run
Cyclone's server side code, while the remaining machines run multithreaded
clients to load the system for our studies. The server side machines are each
equipped with four commodity 82599 10 Gbit ethernet controllers, while each
client machine has one 82599 ethernet controller. All the machines are connected
through the same commodity 10Gbit switch. We use 8 cores on the server machines
for running 8 instance of RAFT, leaving 39 cores to run application code. The
last remaining core is reserved for management code. We use NVML with the non
volatile pool held in DRAM, which proxies for the performance we would expect
from NVDIMMs. We evaluate Cyclone in three stages.

First, we evaluate pure replication performance by replicating RPC calls. There
is crash consistent transaction that remembers the RPC call sequence number -
every RPC call is accepted and once replicated a response is sent to the
client. Figure~\ref{fig:pure_rep} shows the results. We consider two cases, one
where the clients send an RPC call with no payload and one where they add a 512
byte dummy payload. For the no payload case the bottleneck is the message
handling rate. We note that Cyclone can handle small message replication rates
as high as \emph{8 million messages a second}. In the case of RPC calls with 512
byte payloads, the bottleneck is the available network bandwidth. We plot
transmit throughput derived from the message rate as the rate at which data is
sent to replicas from all the 8 quorum leaders and back to the clients.
It is interesting to note that with 3 replicas Cyclone achieves \emph{more than
  40 Gbps} of bandwidth. The reason for this is that the quorum leaders can end
up residing on different machines. Although replication traffic is still limited
by the receive bandwidth the responses to the clients, originating from
different machines have access to independent 4x10GigE NICs, allowing total
transmit throughput to exceed 40 Gbps. Although not discussed in this paper,
Cyclone's control plane is designed to exploit this by trying to spread quorum
leaders across machines.

Also worth noting is that thanks to DPDK's low latency access to the NICs, three
way replication costs as little as 25 us of latency in the unloaded case. This
latency is measured at the client and therefore is four commodity 10 GigE
network hops through the switch, with each hop taking under 7 us including
network stack delays. These results should serve to illustrate that simple
commodity ethernet can indeed provide reasonable replication performance with a
carefully designed software stack that can exploit it.

@@

Next, we consider the effect of adding NVMLs crash consistent
transactions. Figure~\ref{fig:cc_rep} shows the effect of adding a crash
consistent transaction where the application still does nothing but Cyclone
remembers the last sequence number and RPC call result (the application simply
returns the received payload). Throughput is now bottlenecked by NVML allocator
performance at just under a million messages a second. The final result is
adding the hash table in the picture with the clients updating the values of
keys already present in the hash table - a stable workload for easy
comparison. As Figure~\ref{fig:app_rep} shows performance stays at the just
under 1M calls a second mark for pure updates. For comparison we also show a
pure read only case where the clients only query the value of a key - there is
no replication \emph{nor} any crash consistent transaction applied. It is
interesting to note that the pure read only case is bottlenecked at the same
point as the pure replication case of 8 million operations a second.

These results lead us to conclude that Cyclone's provides adequate performance
headroom for replication to NVML applications. In fact our current focus in the
light of these results is towards improving application performance on top of
NVML. NVML is a general purpose library written for a variety of substrates
including traditional SSD based storage and there is adequate room for
improvement of performance for applications running on directly attached
persistent memory with close to DRAM performance.

\begin{figure}
  \includegraphics[width=0.5\textwidth,height=6cm]{results/cc_mops.pdf}
  \caption{Crash Consistency + Replication}
  \label{fig:cc_rep}
\end{figure}

\begin{figure}
  \includegraphics[width=0.5\textwidth,height=6cm]{results/app_mops.pdf}
  \caption{Crash Consistent Replicated Hash Table}
  \label{fig:app_rep}
\end{figure}

\section{Conclusion}
Cyclone allows developers to add availability to applications built on top of
libraries such as NVML and achieve good performance on commodity networking
hardware. A key question is whether the ideas in Cyclone extend to more than
just key-value maps - perhaps to application such as filesystems. We believe
that the requirement for scaling across multiple replication quorums is
can be satisfied by applications where the scalable commutativity rule [scalable
commutativity] applies to common case primitives. Although our intitial release
of Cyclone will be targeted towards key value style applications, our future
research is targeted towards investigating this hypothesis.
\end{document}
